---
title: "Rethinking, Ch2 problems"
author: "Stacey Harmer"
date: "September 17, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2E1
probability of rain on Monday corresponds to: Pr(rain|Monday)
See page 36

## 2E2
Pr(Monday|rain) corresponds to probability of Monday when rainy. or 3.

## 2E3
probability that it is Monday given that it is raining = 1) Pr(Monday|rain)

## 2E4
What does is mean to say the probability of water is 0.7?
At each toss, there is land or water (0 or 1), never a fraction.  But 0.7 tells us something about how often
each of these two outcomes will be observed.

## 2M1
Compute and plot grid approximate posterior distribution for each of hte below, assumign uniform prior.
See R code 2.3

```{r}
library(rethinking)

#1.  W, W, W

# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )


#2.  W, W, W, L

# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )

# compute likelihood at each value in grid
likelihood <- dbinom(3 , size=4 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )


#3.  L, W, W, L, W, W, W

# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )

# compute likelihood at each value in grid
likelihood <- dbinom( 5 , size=7 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )

```


## 2M2
As above, but assume new prior
```{r}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- c(rep( 0 , 10 ), rep(1, 10))

#1.  W, W, W

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )


#2.  W, W, W, L

# compute likelihood at each value in grid
likelihood <- dbinom(3 , size=4 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )


#3.  L, W, W, L, W, W, W

# compute likelihood at each value in grid
likelihood <- dbinom( 5 , size=7 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )

```

## 2M3
Assume 2 globes. Earth, with p = .7.  And Mars, with p = 0 (no water)
One toss of a globe gave 'land'.  what is posterior probability that globe was Earth?
```{r}
# What are the chances that if I tossed Mars, I'd get land?  100%
# what are chances that if I tossed Earth, I'd get land?

likelihood <- dbinom(0, size = 1, prob = .7)
# likelihood is 0.3

# Equally likely to have picked up either globe.  

# posterior = likelihood*prior/avg likelihood

#Pr(Earth|land) = Pr(earth) * Pr(land|earth) / Pr(land)
# Pr(land | Earth)  = 0.3
# Pr(Earth) = 0.5
# Pr(land) = (1+.3)/2 = 0.65

# so:  Pr(earth|land) = 0.5 * 0.3 / 0.65 = 0.2308

```


## 2M4
Assume deck with 3 cards.  each side either B or W.  They are:  B/B, B/W, W/W
So 6 sides. 
One card is picked; B wide up.  What are chances other side is B?
The card drawn was BB or BW.  So 4 sides to consider.  2 ways for BB to have B up.  1 way for WB to have B up.  
So the chosen card twice as likely to be BB as BW.  Thus probaility 2nd side B = 2*BB/(2*BB + 1*BW) = 2/3

## 2M5
4 cards:  BB, BW, WW, BB.  One card chosen, side up is B.  Probability other side will be B?
Need consider 3 cards, or 6 sides.  4 ways for BB to have B side up.  1 way for BW to have B side up.
 so probaility = 4/(4+1) = 0.8
 
## 2M6
Now, B card less likelyt to be pulled.  one each BB, BW, WW cards.  But effectively:
3 WW, 2 BW, and 1 BB
a card is pulled, and B side is up.  What is probaility that other side is B?
effecitvely, we consider 3 cards ( 2 BW, 1 BB) 
2 ways for BB to have B side up; 2 ways for BW to have B side up.  So:  2/(2+2) = 0.5

## 2M7
Treat problem like globe toss. Count all the ways to see each observation for each first card.
first card B:  2 ways for BB to show this, 1 way for BW, 0 way for WW
second card W: 0 ways for BB to show this, 1 way for BW, 2 ways for WW

Table:
first card          second card                total prob
BW = 1 ways B       WW = 2 ways W               1*2 = 2
BB = 2 ways B       WB or WW = 3 ways W         2*3 = 6
WW = 0 ways B       BW = 1 way W                0*1 = 0

So probabilit first card was BB = 6/8 = 0.75

## 2H1
Two panda species.  A gives birth to twins 10% of time.  B gives twins 20% of time
Unknown panda had twins.  What is chance next birth will be twins?

 one way would be:
 If species A, chance of twins would be 10%. 
 If spcies B, chance woudl be 20%.
 So, since we don't know, chance of next birth being twins is 15%

(Should actually be greater than 15% since 1st birth suggests higher chance panda is species B.)

OK:
PR(A|twins) = Pr(twins|A)*Pr(A) / Pr(twins)
= (0.1 * .5)/(0.15) = 0.333  chance that panda was species A
So
Pr(twins|B) * Pr(B) =  0.2 * (1-.33) =  0.134
Pr(twins|A) * Pr(A) = 0.1 * (.33) = 0.033
Sum = 0.134 + 0.033 = 0.167

## 2H2
Assume only 1 birth seen, and it was twins.  Probability that panda is A?
PR(A|twins) = Pr(twins|A)*Pr(A) / Pr(twins)
= (0.1 * .5)/(0.15) = 0.333  chance that panda was species A

## 2H3
first birth twins, second single.  Chance the mmom is A?

1st birth gives us:
PR(A|twins) = Pr(twins|A)*Pr(A) / Pr(twins)
= (0.1 * .5)/(0.15) = 0.333  chance that panda was species A

2nd birth gives us:
PR(A|single) = Pr(single|A)*Pr(A) / Pr(single)
= (0.9 * .33)/(sum(.9*0.33, .8*.666)) = 0.358  chance that panda was species A       


## 2H4
OK, now we have a test.  
It positively identifes A with p of 0.8
It posibitiley idnetified B wiht p of 0.65

It is not clear to me what outcome is if it is wrong.  No ID, or wrong species?
In other words, is there a p of 0.35 that panda is A if test says it is B? 
And a p of 0.2 that it is B if test claims A?    Let's assume so.

Pr(A|pos) = Pr(A) * Pr(pos|A) / Pr(pos)
= (0.5 * .8) /(sum(.8*.5, .35*.5)) = 0.696

OK, now use the birth info as prior (so Pr(A) = 0.358)

Pr(A|pos) = Pr(A) * Pr(pos|A) / Pr(pos)
= (0.358 * .8) /(sum(.8*0.358, .35*(1-0.358))) = 0.56036

