---
title: "Rclub_homework_Dec_12_2016"
author: "Stacey Harmer"
date: "December 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1 : Tomato

Remember the tomato data set generated by Pepe; we first looked at this when we were working on Chapter 9. 
35 accessions for seven species were grown in sun and shade.

Assess whether there is evidence for total length (“totleng”) response to shade and whether this response varies by species. 
Consider whether including accession (“acs”), and/or shelf (“shelf”) using adaptive priors improves the model fit.

Bonus: would it be better to consider shade by accession interactions instead of shade x species?

```{r}
setwd("~/R /R_club/R_Club_2016/Stacey.Harmer_RethinkingHomework/Assigment_2016-12-12")
tom <- read.csv("TomatoR2CSHL.csv")
head(tom)
str(tom)
summary(tom)

library(rethinking)

# get map2stan up and ready
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

```
Make dummy variable, so that sun = 0, shade = 1
```{r}
tom$shade <- ifelse(tom$trt == "L", 1, 0)
head(tom, n=6)
tail(tom, n=20)
```
I want to treat species as a cluster variable.  Do I need to give each species a number?
Try without doing that.

First write model only taking treatment and species into account
After first attempt, learned that I need to trim this dataset. 
Keep accession (acs) and shelf and totleng and treatment and speices but ditch the rest

```{r}

tom.trim <- tom[, c(1,5, 15, 24,26)]
   
head(tom.trim)
str(tom.trim)

tom.trim$sp_index <- coerce_index(tom.trim$species)

```


```{r}
mTom.1 <- map2stan(
  alist(
    totleng ~ dnorm(mu, sigma),
    mu <- a_species[sp_index] + b_species[sp_index]*shade,
    a_species[sp_index] ~ dnorm(0, 10),
    b_species[sp_index] ~ dnorm(0,10),
    sigma ~ dnorm(0, 10)), data=tom.trim, iter=5000, warmup=2000, chains =2)
plot(mTom.1)
precis(mTom.1, depth=2)
plot(precis(mTom.1, depth = 2))
```

Consider whether including accession (“acs”), and/or shelf (“shelf”) using adaptive priors improves the model fit.

```{r}
# OK, include accession and then shelf; first make them an index

tom.trim$acs_index <- coerce_index(tom.trim$acs)
tom.trim$shelf_index <- coerce_index(tom.trim$shelf)
head(tom.trim)

# include accession, using adaptive prior
mTom.2 <- map2stan(
  alist(
    totleng ~ dnorm(mu, sigma),
    mu <- a_species[sp_index] + a_acs[acs_index] + b_species[sp_index]*shade,
    a_species[sp_index] ~ dnorm(0, 10),
    b_species[sp_index] ~ dnorm(0,10),
    a_acs[acs_index] ~ dnorm(0, sigma_acs),
    sigma_acs ~ dnorm(0, 10),
    sigma ~ dnorm(0, 10)), data=tom.trim, iter=5000, warmup=2000, chains =2)

plot(mTom.2)
# neff not great for the a_acs values
precis(mTom.2, depth=2)
par(mfrow = c(1,1))
plot(precis(mTom.2, depth = 2))
# looks a bit worse than the simpler model

# include shelf, using adaptive prior
mTom.3 <- map2stan(
  alist(
    totleng ~ dnorm(mu, sigma),
    mu <- a_species[sp_index] + a_shelf[shelf_index] + b_species[sp_index]*shade,
    a_species[sp_index] ~ dnorm(0, 10),
    b_species[sp_index] ~ dnorm(0,10),
    a_shelf[shelf_index] ~ dnorm(0, sigma_acs),
    sigma_acs ~ dnorm(0, 10),
    sigma ~ dnorm(0, 10)), data=tom.trim, iter=5000, warmup=2000, chains =2)

plot(mTom.3)
precis(mTom.3, depth=2) # again, n-eff not great
par(mfrow = c(1,1))
plot(precis(mTom.3, depth = 2))
# looks like shlef has a bigger effect than species
```
now include both acs and shelf in the model
```{r}
mTom.4 <- map2stan(
  alist(
    totleng ~ dnorm(mu, sigma),
    mu <- a_species[sp_index] + a_shelf[shelf_index] + a_acs[acs_index] + b_species[sp_index]*shade,
    a_species[sp_index] ~ dnorm(0, 10),
    b_species[sp_index] ~ dnorm(0,10),
    a_acs[acs_index] ~ dnorm(0, sigma_acs),
    a_shelf[shelf_index] ~ dnorm(0, sigma_acs),
    sigma_acs ~ dnorm(0, 10),
    sigma_acs ~ dnorm(0, 10),
    sigma ~ dnorm(0, 10)), data=tom.trim, iter=5000, warmup=2000, chains =2)
# Julin: set alpha to be about the mean;
# set sigma to use dcauchy distribution.  
#  any maybe center the beta on a positive number also; perhaps 20?  

plot(mTom.4) 
precis(mTom.4, depth=2) # n-eff better now
par(mfrow = c(1,1))
plot(precis(mTom.4, depth = 2))

```
compare these 4 models

```{r}
compare(mTom.1, mTom.2, mTom.3, mTom.4)

```
As I suspected based on the n_eff values, model 4 is much preferred.  

NOTE FROM CLASS - need to think about the priors harder!  Should have made prior for
shade treatment positive (say, 20) and let the shelf effect prior be 0.  

LOOK AT PARAMETER estimates as well - does it makes sense that shelf
has larger effect than species ?!?!?!

## Problem 2: Smoking deaths among doctors

In 1961 Doll and Hill sent out a questionnaire to all men on the British Medical Register inquiring about their smoking habits. Almost 70% of such men replied. Death certificates were obtained for medical practitioners and causes of death were assigned on the basis of these certificates. The breslow data set contains the person-years of observations and deaths from coronary artery disease accumulated during the first ten years of the study.

Analyse this data set to determine the posterior probability that smoking increases death by coronary artery disease, that age increases death by coronary artery disease, and that there is an interaction between age and smoking.

You can load the data set and learn about the columns using the commands below

data("breslow",package = "boot")
help("breslow",package ="boot")
breslow
You can think of “person-years” as the number of observations

Note: You almost certainly have the boot package on your computer, but if you do not have the boot package on your computer then you will need to install.packages("boot")

Note: do NOT do library(boot). This will make the logit function from boot over-ride the one from rethinking
Note: You will probably need to do:

breslow$n <- as.integer(breslow$n)

before you can analyze the problem 2 data

```{r}

data("breslow",package = "boot")
help("breslow",package ="boot")
breslow
breslow$n <- as.integer(breslow$n)
str(breslow)
```

Analyse this data set to determine the posterior probability that smoking increases death by coronary artery disease
y is the number of deaths; I think I want to convert to proportions
so number of deaths (y/n)

Notes in class: NOPE, I didn't need to do that.

```{r}
breslow$fxn_death <- breslow$y/breslow$n

mDoc.1 <- map2stan(
  alist(
    y ~ dbinom(n, p),
    logit(p) <- a + beta_smoke*smoke,
    a ~ dnorm(0, 5),
    beta_smoke ~ dnorm(0, 5)), 
  data=breslow, iter=5000, warmup=2000, chains =2)

plot(mDoc.1)
precis(mDoc.1, depth=2)
par(mfrow=c(1,1))
plot(precis(mDoc.1, depth = 2))

```
Note the above numbers are on teh log odds score
```{r}
logistic(-5.96)
#  0.002573273 prob of dying if not a smoker

logistic(-5.96 + .55)
# 0.004451734 prob of dying if you smoke
# difference
0.004451734/0.002573273 # = 1.73  so almost twice as likely to die of CV disease if you smoke  (this is per year)

```

Analyze data to see if age increases death by coronary artery disease

age : Julin made it an index

```{r}
breslow$age_id <- coerce_index(breslow$age)

mDoc.2 <- map2stan(
  alist(
    y ~ dbinom(n, p),
    logit(p) <- alpha_age[age_id],
    alpha_age[age_id] ~ dnorm(0, 5) ), 
  data=breslow, iter=5000, warmup=2000, chains =2)

plot(mDoc.2)
precis(mDoc.2, depth=2)
par(mfrow=c(1,1))
plot(precis(mDoc.2, depth = 2)) 

```

and that there is an interaction between age and smoking.
```{r}

```




##Problem 3: Cane Sugar

This data comes from an experiment to measure disease resistance in different varieties of sugar cane.

Is there evidence of differences in disease resistance in the different varieties? Does including an adaptive prior for Block improve the model fit?

You can get the data and learn about it with:

data("cane",package="boot")
help("cane",package="boot")
head(cane)
summary(cane)

