---
title: "Ch 13 homework, last section"
author: "Stacey Harmer"
date: "February 9, 2017"
output: 
  html_document: 
    keep_md: yes
---

#13M4
use WAIC to compare Gaussian process modeling of Oceanic tools to models fit for same 
data in chapter 10.  Think about effective # of parameters

First, models from Ch 10.

```{r}

library(rethinking) 
data(Kline)
d.Kline <- Kline
d.Kline


```
model total tools as output
```{r}
d.Kline$log_pop <- log(d$population)
d.Kline$contact_high <- ifelse( d$contact=="high" , 1 , 0 )

```
model includes interaction between log_pop and contact rate
WHY IS POISSON A GOOD MODEL IN THE FIRST PLACE?

```{r}
m10.10 <- map( 
alist(
total_tools ~ dpois( lambda ),
log(lambda) <- a + bp*log_pop +
bc*contact_high + bpc*contact_high*log_pop,
a ~ dnorm(0,100),
c(bp,bc,bpc) ~ dnorm(0,1)
),
data=d.Kline )
```

now the other models
First, no interaction
```{r}
# no interaction 10.45
m10.11 <- map(
alist(
total_tools ~ dpois( lambda ),
log(lambda) <- a + bp*log_pop + bc*contact_high,
a ~ dnorm(0,100),
c(bp,bc) ~ dnorm( 0 , 1 )), data=d.Kline )

```
 then, no contact rate or no log pop
 
 
```{r}
# no contact rate 
m10.12 <- map(
alist(
total_tools ~ dpois( lambda ),
log(lambda) <- a + bp*log_pop,
a ~ dnorm(0,100),
bp ~ dnorm( 0 , 1 )), data=d.Kline)

# no log-population
m10.13 <- map(
alist(
total_tools ~ dpois( lambda ),
log(lambda) <- a + bc*contact_high,
a ~ dnorm(0,100),
bc ~ dnorm( 0 , 1 )), data=d.Kline )
```

```{r}
# intercept only 
m10.14 <- map(
alist(
total_tools ~ dpois( lambda ),
log(lambda) <- a,
a ~ dnorm(0,100)), data=d.Kline)

# compare all using WAIC
# adding n=1e4 for more stable WAIC estimates
# will also plot the comparison
( islands.compare <- compare(m10.10,m10.11,m10.12,m10.13,m10.14,n=1e4) )
plot(islands.compare)
```

And the model from Ch 13
```{r}

```
 Use the WAIC function to compare the models
 
 
 
 # next problem
 ?Oxboys
 ??dist
 
