---
title: "Ch12_problems_2016-11-11.Rd"
author: "Stacey Harmer"
date: "November 11, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 12.M3
Frog survival model; comparing Gaussian and Cauchy distributions for the varying intercepts

First, the model from the book with Gaussian distributions for intercepts

```{r,  results = 'hide' }
library(rethinking)
data("reedfrogs")
??reedfrogs
d <- reedfrogs
str(d)

# get map2stan up and ready
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# make the tank cluster variable
d$tank <- 1:nrow(d)
head(d)


m12.M3.1 <- map2stan( 
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1)
  ), data=d , iter=4000 , chains=4 )

```

Next, the same but Cauchy intstead of normal

p 260: The Cauchy distributions in the model definitions are implicitly half-Cauchy, a Cauchy defined over
the positive reals only. This is because they are applied to a parameter, usually sigma , that is strictly positive.
Stan figures out that you meant for it to be half-Cauchy.  SO I guess I just won't specify??
```{r,results = 'hide'  }
m12.M3.2 <- map2stan( 
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dcauchy( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1)
  ), data=d , iter=4000 , chains=4 )

```

Compare the posterior means of the intercepts, a-tank, to the posterior means produced in the chapter,
using the customary Gaussian prior. Can you explain the pattern of differences?
```{r }
precis(m12.M3.2,depth=2) # depth=2 displays varying effects
plot(precis(m12.M3.1, depth =2))
plot(precis(m12.M3.2,depth=2)) # also plot

par(mfrow = c(1,2))
plot(precis(m12.M3.1, depth =2)) # gaussian 
plot(precis(m12.M3.2,depth=2)) # Cauchy
```

 Overall, much lower SD with cauchy (although some are really uncertain, like 2, 7, 20, 38)
These large SD tanks had 100% survival.  
"at any moment in a Cauchy sampling process, you are able to draw an extreme value that overwhelms all the previous draws "
I guess that is what is going on - with a high prop survival, can get very wrong estimate.


## Question 12.H1
Fr