---
title: "Ch10_problems_2016-10-21"
author: "Stacey Harmer"
date: "10/21/2016"
output: 
  html_document: 
    keep_md: yes
---

Read the rest of chapter 10  (maybe later)

Do problem 10H3

First, look at the eagle dataset
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
?eagles
library(MASS)
data(eagles)
eagles
```
y is number os successful attempts, n total number of attempts, P size of pirate, A age of pirate, V size of victim

Build a binomial GLM of successful pirating attempts
First need to turn P, A, and V into dummy variables.
For P, 1 = large, 0 = small
For A, 1 = adult, 0 = immature
For V, 1 = large, 0 = small

```{r}
eagles$Pd <- ifelse(eagles$P == "L", 1, 0)
eagles$Ad <- ifelse(eagles$A == "A", 1, 0)
eagles$Vd <- ifelse(eagles$V == "L", 1, 0)
```
Now construct a binomial GLM; I'm modeling the successful attempts (n)

```{r}
eagle.glm <- map(
  alist(
    y ~ dbinom(n, p) ,
    logit(p) <- a + bp*Pd + ba*Ad + bv*Vd ,
    a ~ dnorm(0,10),
    c(bp, ba, bv) ~ dnorm(0, 5)
  ),
data = eagles)
precis(eagle.glm)
pairs(eagle.glm)
plot(precis(eagle.glm))
# looks like b.p. is sig positive, b.v sig negative, and b.a slightly positive
```
Now do this using map2stan

```{r}
# get map2stan up and ready
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

eagle.stan <- map2stan( eagle.glm , chains = 4, data=eagles , iter=1e4 , warmup=1000 )
precis(eagle.stan) # looks fine
plot(eagle.stan)

par( mfrow = c( 1,1 ) )
plot(precis(eagle.stan)) # similar to MCMC
pairs(eagle.stan) # ditto
compare(eagle.glm, eagle.stan)

```
These plots look very similar, so quadratic approx seemed OK

ACTUALLY, pairs plots in STAN are asymmetric - should have switched to STAN for below

Plot posterior predictions.  Compute and display: 1) predicted _prob_ of success + its 89% interval for each row
predicted success _count_ and its 89% interval

First make dummy data for each type of interaction (between P, A, and V)
```{r}
eagle.pred <- data.frame(
  Pd= c(1,1,1,1,0,0,0,0),
  Ad= c(1,1,0,0,1,1,0,0),
  Vd=rep(c(1,0), 4))
```

Build prediction ensemble
```{r}
eagle.ensemble <- ensemble(eagle.glm, data = eagles) # probably better way for 1 model
# YES< COULD USE LINK function
# summarize
pred.eagle <- apply(eagle.ensemble$link, 2, mean)
pred.eagle.PI <- apply(eagle.ensemble$link, 2, PI)
# these are probabilities
```
Now I will plot the data plus the predictions

```{r}
# empty plot frame with good axes 
plot( 0 , 0 , type="n" , xlab="pirate_size/pirate_age/victim_size" ,
ylab="proportion success" , ylim=c(0,1) , xaxt="n" ,
xlim=c(1,8) )
axis( 1 , at=1:8 , labels=c("1/1/1","1/1/0","1/0/1","1/0/0", "0/1/1", "0/1/0", "0/0/1", "0/0/0") )

# I think I need to determine actual prob of success in the dataset
eagles$prob <- eagles$y/eagles$n
# now plot these data
points(eagles$prob)
# and now add my posterior predictions
points(pred.eagle, col = "blue")
# now try to add the PI

for ( i in 1:8 ) {
  x <- i
  y1 <- pred.eagle.PI["5%", x]
  y2 <- pred.eagle.PI["94%", x]
  lines(  c(y1,y2) , col=rangi2 , lwd=2 )
} # nope, that was wrong

for ( i in 1:8 ) {
  x <- i
  y1 <- pred.eagle.PI["5%", x]
  y2 <- pred.eagle.PI["94%", x]
  lines(  c(y1,y2), c(x,x),  lwd=2)
}

```
Not quite right with the graphing.  But overall, the posterior look pretty good
```{r}
#how about:
postcheck(eagle.glm, n=1e4)  #whoa!  probabilities and ranges are all there.  Pretty easy! 
# outlier is the last one (small young pirate, small victim; actual success higher than predicted)
par( mfrow = c( 1,1 ) )
postcheck(eagle.stan, n=1e4)
```

OK, how do I graph out predicted success counts?

SHOULD MULTIPLY PROBABILITY BY ACTUAL N IN THE REAL DATA FRAME

JULIN MADE model with age*size as a factor.  This is a negative term in model.
if both old and large, then less likely by a bit. Kind of negates the age.
So size is the main thing, seems like.





Do the following problem:

The attached data sheet is from Liron and records germination for M82 and pen at two temperatures.

Perform a Bayesian analysis to determine whether or not the two species have different germination rates, whether temperature affects germination, and whether there is a temperature by germination interaction.  Determine the posterior estimates for germination rate for each species at each temperature and determine the confidence intervals around these estimates.

A couple of hints: if you are using Stan you will need to remove the "." from any variable names.

could use gsub to remove dots from column names

You may want to baseline temperature so that the lowest value is 0 rather than 14
 may make temperature a dummy variable instead

You may want to make a dummy variable for species.  Or alternatively use multiple intercepts.
  make M82 zero and Penn 1  

