---
title: "Chapter-05-part2-assignment"
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Stacey__


# For 04/25/2016

## 5M2
***
Invent your own example of a masked relationship. An outcome variable should be correlated
with both predictor variables, but in opposite directions. And the two predictor variables should be
correlated with one another  

cycling speed = outcome variable; predictor.1 = wind resistance and predictor.2 = rpm (pedaling speed)  


## 5H1
##### Fit two bivariate Gaussian regressions, using map: (1) body weight as a linear function of territory
##### size (area), and (2) body weight as a linear function of groupsize. Plot the results of these
##### regressions, displaying the MAP regression line and the 95% interval of the mean. Is either variable
##### important for predicting fox body weight?

```{r}
library(rethinking)
data("foxes")
foxes <- foxes
summary(foxes)

# 1.  body weight as linear function of territory size (area)

m.area <- map( 
              alist(
                weight ~ dnorm(mu,sigma),
                mu <- a + b*area  ,
                a ~ dnorm(3,10),  # not sure what reasonable prior is; will try this
                b ~ dnorm(0,10),
                sigma ~ dunif(0,10)
              ),
              data=foxes )
precis(m.area)
plot(precis(m.area))  # b not very useful

area.seq <- seq(0,6, length.out = 100)
mu.area <- link(m.area, data=data.frame(area=area.seq))

mu.area.HPDI <- apply(mu.area, 2, HPDI, prob=0.95)

plot(weight ~ area, data=foxes, col= rangi2)
abline(m.area)

shade(mu.area.HPDI, area.seq) # not much data within 95% interval

# 2.  body weight as linear function of groupsize

m.size <- map( 
              alist(
                weight ~ dnorm(mu,sigma),
                mu <- a + b*groupsize  ,
                a ~ dnorm(4,10),  # not sure what reasonable prior is; will try this
                b ~ dnorm(0,10),
                sigma ~ dunif(0,10)
              ),
              data=foxes )
precis(m.size) # the bigger the group, the lower the weight
plot(precis(m.size)) # but effect is quite small

groupsize.seq <- seq(0,10, length.out = 100)
mu.groupsize <- link(m.size, data=data.frame(groupsize=groupsize.seq))

mu.groupsize.HPDI <- apply(mu.groupsize, 2, HPDI, prob=0.95)

plot(weight ~ groupsize, data=foxes, col= rangi2)
abline(m.size)

shade(mu.groupsize.HPDI, groupsize.seq) # not much data within 95% interval
```

## 5H2
##### Now fit a multiple linear regression with weight as the outcome and both area and groupsize
##### as predictor variables. Plot the predictions of the model for each predictor, holding the other predictor
##### constant at its mean. What does this model say about the importance of each variable? Why do you
##### get different results than you got in the exercise just above?
```{r}

library(rethinking)
data("foxes")
foxes <- foxes
summary(foxes)

#   body weight as linear function of area and groupsize

m.area.size <- map( 
              alist(
                weight ~ dnorm(mu,sigma),
                mu <- a + b.a*area   + b.g*groupsize,
                a ~ dnorm(3,10),  # not sure what reasonable prior is; will try this
                b.a ~ dnorm(0,10),
                b.g ~ dnorm(0,10),
                sigma ~ dunif(0,10)
              ),
              data=foxes )
precis(m.area.size) 
plot(precis(m.area.size))
# now see that area has positive, groupsize negative effect on weight
# small but non-zero effects

# plot predictions (counterfactual plots)
# first, vary groupsize
area.mean <- mean(foxes$area)
groupsize.seq <- seq(0,10, length.out = 100)
pred.data.groupsize <- data.frame(groupsize=groupsize.seq, area = area.mean)

# use these data to get mean predicted weights.

mu.weight.pred.groupsize <- link(m.area.size, data=pred.data.groupsize)
mu.weight.pred.groupsize.mean <- apply(mu.weight.pred.groupsize, 2, mean)
mu.weight.pred.groupsize.PI <- apply(mu.weight.pred.groupsize, 2, PI, prob=0.95)

# simulate weight outcomes of model
weight.sim <- sim(m.area.size, data=pred.data.groupsize, n=1e4)
weight.PI <- apply(weight.sim, 2 ,PI)

# display predictions, hiding raw data
plot(weight ~ groupsize, data=foxes, type = "n")
mtext("area = 0")
lines(groupsize.seq, mu.weight.pred.groupsize.mean)
shade(mu.weight.pred.groupsize.PI, groupsize.seq)
shade(weight.PI, groupsize.seq)


# Now, vary area (leaving groupsize unaltered)
groupsize.mean <- mean(foxes$groupsize)
area.seq <- seq(0,10, length.out = 100)
pred.data.area <- data.frame(groupsize=groupsize.mean, area = area.seq)

# use these data to get mean predicted weights.

mu.weight.pred.area <- link(m.area.size, data=pred.data.area)
mu.weight.pred.area.mean <- apply(mu.weight.pred.area, 2, mean)
mu.weight.pred.area.PI <- apply(mu.weight.pred.area, 2, PI, prob=0.95)

# simulate weight outcomes of model
weight.sim.2 <- sim(m.area.size, data=pred.data.area, n=1e4)
weight.PI.2 <- apply(weight.sim.2, 2 ,PI)

# display predictions, hiding raw data
plot(weight ~ area, data=foxes, type = "n")
mtext("area = 0")
lines(area.seq, mu.weight.pred.area.mean)
shade(mu.weight.pred.area.PI, area.seq)
shade(weight.PI.2, area.seq)
```
  
  ## 5H3
##### Finally, consider the avgfood variable. Fit two more multiple regressions: (1) body weight
##### as an additive function of avgfood and groupsize, and (2) body weight as an additive function of
##### all three variables, avgfood and groupsize and area. Compare the results of these models to the
##### previous models you’ve fit, in the first two exercises. 


```{r}
# 1.  weight as function of avgfood and groupsize
m.food.group <- map( 
              alist(
                weight ~ dnorm(mu,sigma),
                mu <- a + b.g*groupsize + b.f*avgfood ,
                a ~ dnorm(4,10),  # not sure what reasonable prior is; will try this
                b.g ~ dnorm(0,10),
                b.f ~ dnorm(0,10),
                sigma ~ dunif(0,10)
              ),
              data=foxes )
precis(m.food.group) # the bigger the group, the lower the weight; the more food, the greater the weight
plot(precis(m.food.group))  # really large SE for food effect

# let's look at the gt beween teh model and hte actual data
mu.food.group <- link(m.food.group)
# summarize samlpes across the cases
mu.food.group.mean <- apply(mu.food.group, 2, mean)
mu.food.group.PI <- apply(mu.food.group, 2, PI, prob=0.95)

# simulate observations, starting from real data
weight.sim.1 <- sim(m.food.group, n = 1e4)
weight.PI.1 <- apply(weight.sim.1, 2, PI, prob = 0.95)

# now plot predictions against observed
plot(foxes$weight ~ mu.food.group.mean  , col = rangi2,
     xlab = "observed weight", ylab= "predicted weight")
abline(a =0, b=1, lty=2)
for (i in 1:nrow(foxes) )
     lines( rep(foxes$weight[i],2) , c(mu.food.group.PI[1, i], mu.food.group.PI[2, i]),
            col=rangi2)

# I messed up these lines. . .

# try computing residuals instead

weight.food.group.resid <- foxes$weight - mu.food.group.mean
# get ordering by weight
w <- order(weight.food.group.resid)
# make the plot
dotchart( weight.food.group.resid[w] , labels=foxes$group[w] ,  cex=0.6 )
abline( v=0 , col=col.alpha("black",0.2) )
for ( i in 1:nrow(foxes) ) {
j <- w[i] # which group in order
lines( foxes$weight[j]-c(mu.food.group.PI[1,j],mu.food.group.PI[2,j]) , rep(i,2) )
points( foxes$weight[j]-c(weight.PI.1[1,j],weight.PI.1[2,j]) , rep(i,2),
pch=3 , cex=0.6 , col="gray" )
}

# better

```

##### (a) Is avgfood or area a better predictor of body weight?

```{r}
# 1.  as above, but replace avgfood with area
m.area.group <- map( 
              alist(
                weight ~ dnorm(mu,sigma),
                mu <- a + b.g*groupsize + b.a*area ,
                a ~ dnorm(4,10),  # not sure what reasonable prior is; will try this
                b.g ~ dnorm(0,10),
                b.a ~ dnorm(0,10),
                sigma ~ dunif(0,10)
              ),
              data=foxes )
precis(m.area.group) # the bigger the group, the lower the weight; the more food, the greater the weight
plot(precis(m.area.group))  # area has smaller effect than avgfood, but SE much smaller.  This is likely better


```

#####  If you had to choose one or the other to include in a model, which would it be? Support your
##### assessment with any tables or plots you choose. (b) When both avgfood or area are in the same
##### model, their effects are reduced (closer to zero) and their standard errors are larger than when they
##### are included in separate models. Can you explain this result?
***
no time to run it, but likely this is because larger area is highly correlated with avgfood

```{r}
head(foxes)
pairs(foxes)
pairs( ~ avgfood + area  ,
   data=foxes , col=rangi2 )
# these look highly correlated
cor(foxes$avgfood, foxes$area)
# 0.8831038
```
yes, these values are highly positively corelated.