---
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Stacey:__

# For 04/11/2016

```{r}
library(rethinking)
```


## 5E1
#####  5E1. Which of the linear models below are multiple linear regressions?
#####  (2) mu.i = b.x * x.i + b.z * z.i
#####  (4)  mu.i = a + b.x*x.i + b.z * z.i

#####   (I dont' think 3 qualifies as only one beta here)

## 5E2.   Write down a multiple regression to evaluate the claim: 
## Animal diversity is linearly related to latitude, but only after controlling for plant diversity. 

#####   (  I think it would be:  mu.d = a + b.l * lat + b.p * pl.d
#####   ( where mu.d is mean animal diversity
#####   ( a is intercept
#####   ( b.l is the relationship between latitude and diversity
#####   ( lat is latitude
#####   ( b.p is relatinship between plant divserity and animal diversity
#####   ( pl.d is plant diverstiy


## 5M1.  Invent your own example of a spurious correlation. 
#####  An outcome variable should be correlated with both predictor variables. 
#####  But when both predictors are entered in the same model, 
#####  the correlation between the outcome and one of the predictors 
#####   should mostly vanish (or at least be greatly reduced).

##### well, this is spelled out on p. 135
```{r}
n <- 200
real.x <- runif(n, min = 10, max = 20)
spur.x <- rnorm(n, real.x)
my.y <- rnorm(n, real.x)
my.d <- data.frame(my.y, real.x, spur.x)
pairs(my.d)
summary(my.d)

# can I make a linear regression predicting y?
library(rethinking)
my.M1.model <- map(
  alist(
    my.y ~ dnorm (mu, sigma) ,
    mu <- a + bS * spur.x + bR * real.x,
    a ~ dnorm(10, 10) ,
    bS ~ dnorm(0 , 1) ,
    bR ~ dnorm(0 , 1) ,
    sigma ~ dunif(0, 10)
  ), data = my.d
)
precis(my.M1.model)
plot(precis(my.M1.model))
#  bS is close to 0, while bR is close to 1.  OK!
```

## 5M3  Consider how a high divorce rate might cause a higher marriage rate. 
##  Can you think of a way to evaluate this relationship, using multiple regression?

##### could imagine that recently divorced people would be eager to marry again.
##### evaulation with multiple regression: As in Fig 4.5?  (p 103)
##### I think I'd model using map function, model being
##### mu <- a + b*divorce.rate
##### then I'd extract samples from the posterior and plot relative to real data
##### as shown on page 102


## 5M4
### Find a list of LDS population by State and use those numbers as a predictor variable, predicting divorce
### rate using marriage rate, median age at marriage, and percent LDS population (possibly standardized).
### You may want to consider transformations of the raw percent LDS variable.

#### found appropriate table on Wikipedia; saved to my hard drive

#### what does he mean by standardized percentile?  Ah, maybe so that it's fraction mormons in US by state?

```{r}
setwd("~/R /R_club/R_Club_2016/Ch5_2016-04-22")
mormons <- read.csv("mormons_by_state.csv")
head(mormons)
tail(mormons, n=10)
summary(mormons)
sum(mormons$percent.LDS)
mormons$percent.std <- mormons$percent.LDS / sum(mormons$percent.LDS)  

# standardize LDS predictor
mormons$percent.LDS.s <- (mormons$percent.LDS - mean(mormons$percent.LDS))/
  sd(mormons$percent.LDS)
dim(mormons)

# now simplify and merge info
mormons.LDS.s <- mormons[, c(1,14)]
head(mormons.LDS.s)

data("WaffleDivorce")
divorce <- WaffleDivorce
head(divorce)
divorce <- merge(divorce, mormons.LDS.s, by.x = "Location", by.y = "State")
head(divorce)

divorce$Marriage.s <- (divorce$Marriage - mean(divorce$Marriage))/sd(divorce$Marriage)
divorce$MedianAgeMarriage.s <- (divorce$MedianAgeMarriage-mean(divorce$MedianAgeMarriage))/sd(divorce$MedianAgeMarriage)

### now predict divorce rate with all parameters

m5M4.problem <- map(
  alist(
    Divorce ~ dnorm (mu, sigma) ,
    mu <- a + bR * Marriage.s + bA * MedianAgeMarriage.s + bM * percent.LDS.s,
    a ~ dnorm(10, 10) ,
    bR ~ dnorm(0 , 1) ,
    bA ~ dnorm(0 , 1) ,
    bM ~ dnorm(0 , 1) ,  # just like others .. .
    sigma ~ dunif(0, 10)
  ), data = divorce
)
precis(m5M4.problem)

plot(precis(m5M4.problem))
# Age and LDS both predict less divorce
```