---
title: "brms_intro_homework"
author: "Stacey Harmer"
date: "January 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(rethinking)

# get map2stan up and ready
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

library(brms)
library(ggplot2)
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Question 1 
Fit model 10.9 (R code 10.28) with brms; compare to map2stan version.
 Test whether the coefficient for gender is different from 0 in the brms model.
This is hte grad school admission data (see page 304)
male vs female admissions.  

```{r cars}
data("UCBadmit")
d <- UCBadmit
d
# make male numeric
d$male <- ifelse( d$applicant.gender=="male" , 1 , 0 ) 
d

# make index for department
d$dept_id <- coerce_index( d$dept )

# model with unique intercept for each dept
# model with male difference as well
m10.9 <- map(
  alist(
    admit ~ dbinom( applications , p ) ,
    logit(p) <- a[dept_id] + bm*male ,
    a[dept_id] ~ dnorm(0,10) ,
    bm ~ dnorm(0,10)
  ) , data=d )
summary(m10.9)
precis(m10.9, depth = 2)
```

Now make same model with brms
```{r}
d
m10.9.brms <- brm(admit | trials(applications) ~ dept + male,
                  family = "binomial",
                  prior=set_prior("normal(1,10)", class = "b"),
                  data = d
                  )
summary(m10.9.brms)
plot(m10.9.brms, ask = FALSE)

```

Compare the two models
```{r}
waic(m10.9, m10.9.brms)
# doesn't work.  not compare, not waic, not loo
# but visual inspection of the parameter estimates shows they are quite similar
precis(m10.9, depth = 2)
summary(m10.9.brms)

```

Now see if gender coeff is really different from 0

```{r}
hypothesis(m10.9.brms, "male = 0")
plot(hypothesis(m10.9.brms, "male = 0"))
# it could easily be zero.
```

# Question 2

Refit models 12.1 and 12.2 (Rcode 12.2 and 12.3) with brms

Tadpole data.  See page 357.  Multilevel model, estimating boht intercept for each tanks
and variation among tanks
```{r}
data("reedfrogs")
d <- reedfrogs
head(d)
tail(d)
```

 first, for model 12.1
 
 The map2stan model is:
 m12.1 <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dnorm( 0 , 5 )
  ),
  data=d )

So now make a brms model; this again is binomial.  Use a 0 to indicate separate intercept for each level
```{r}
d$tank <- 1:nrow(d)
d$tank <- as.factor(d$tank) # didn't work when tank was a number
head(d)  
m12.1.brms <- brm(surv | trials(density) ~ 0 + tank, 
                  family = "binomial",
                  prior = set_prior("normal(0,5)", class = "b"),
                  data = d
                  )
summary(m12.1.brms)
plot(m12.1.brms, ask = FALSE) #  one b value per row

```

And now for 12.2; here, want to have priors both for tank and for the SD of tank

m12.2 <- map2stan( 
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1)
  ), data=d , iter=4000 , chains=4 )

Transforms into random effect model now in brms
```{r}
m12.2.brms <- brm(surv | trials(density) ~ 0 + (1|tank), 
                  family = "binomial",
                  prior = c(set_prior("normal(0,1)", class = "Intercept"),
                  set_prior("cauchy(0,1)", class = "sd")), #half cauchy prior for sd
                  data = d
                  )
summary(m12.2.brms)
plot(m12.2.brms, ask = FALSE)
# does tank have to be a number now?

d$tank <- as.numeric(d$tank)  # then re-run the above model
summary(m12.2.brms)
plot(m12.2.brms, ask = FALSE)
# appears to have the tank esitmates but it doesn't report them to me
?ranef  # to extract random effects
ranef(m12.2.brms) # there they are.  Doesn't care if tank is a factor or a number
```

# Question 3
Use both brms and map2stan to fit a model for intleng as a function of species, trt and their interaction, and include shelf as a random effect.
