---
title: "Tomato hy problem, Ch9"
author: "Stacey Harmer"
date: "August 23, 2016"
output: 
  html_document: 
    keep_md: yes
---

Do a Bayesian analysis of hypocotyl length (hyp) in the attached data sheet.

1) Consider treatment effects (trt) species effects (species) and their interaction.  What is the best model given these possible predictors?

2) Use "S. chmielewskii" as the reference.  For each of the other species evaluate the hypothesis that their response to shade differs from S. chmielewskii.

## ignore the interaction term for now.

1) Does the best model include species, trt, or both?

2) Evaluate the hypothesis that trt has an effect on hypocotyl length

## new instructions
2) Redo the tomato hyp anaysis (or at least the final model with species and treatment) but using square root transformation.  Are model fits improved?  Does it change the parameter estimates?

3) (optional) fit a model with a species by treatment interaction.  Which species have a trt response that is different from S. chm?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)

# look at dataset
setwd("~/R /R_club/R_Club_2016/Stacey.Harmer_RethinkingHomework/Assignment_Chapter_09")
data <- read.csv("TomatoR2CSHL.csv")
head(data)
summary(data) # note the NAs
dim(data) # 1008   25

## just consder hyp; get rid of internode and leaf data; retain only trt and species (discard shelf, who, etc)
hy.data <- data[, c(6,9, 24)]
head(hy.data)
summary(hy.data)  # now I don't have any NAs

# as instructed, I'll perform square root transformation
hy.data$sqrt.hy <- sqrt(data$hy)

```

Julin first inspected the data in various ways; I will do that as well
```{r}
library(ggplot2)
library(reshape2)

# density plot
pl <- ggplot(data = hy.data, aes(x=hyp, fill= trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl  # nice.  

# box plots
pl <- ggplot(data=hy.data,aes(y=hyp,x=species,fill=trt))
pl <- pl + geom_boxplot()
pl # also nice

# now the same for square root transformed

pl <- ggplot(data = hy.data, aes(x=sqrt.hy, fill= trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl + ggtitle("square root transformed") # also looks good

# and now for log2 transformed

pl <- ggplot(data = hy.data, aes(x=log2(hyp), fill= trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl + ggtitle("log2 transformed")

# not clear that square root is 'best'

# normality tests:
by(hy.data$hyp,list(hy.data$trt,hy.data$species),shapiro.test)
by(sqrt(hy.data$hyp),list(hy.data$trt,hy.data$species),shapiro.test)
by(log2(hy.data$hyp),list(hy.data$trt,hy.data$species),shapiro.test)

# ok, I see that the 'W' values are highest for sqrt

# now make trt a categorical variable
levels(hy.data$trt)
hy.data$trt2 <- as.numeric(hy.data$trt)-1  # H is 0, L is 1
summary(hy.data)

#categorical variables for species
hy.data$index <- 1:nrow(hy.data)  # ensures all rows are retained
head(hy.data)
hy.data2 <- dcast(hy.data,index + hyp + trt2 ~ species, value.var="species",fun.aggregate=length)

head(hy.data2)

# how to make a model?  I guess using maptostan

#first, model treatment only.  So get rid of excess data.

data.trt <- hy.data2[, c(2:3)]
head(data.trt)

m.trt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt2,
    a ~ dnorm(mu=33,sd=10),
    bT ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.trt,chains = 4,cores = 1)

plot(m.trt)  # chains look fine
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.trt)
precis(m.trt)
plot(precis(m.trt))

```

Now model the effect of species only
```{r}
data.species <- hy.data[,c("hyp","species")]
data.species$id <- 1:nrow(data.species)
head(data.species)
data.species <- dcast(data.species, hyp + id ~ species, value.var="species",   fun.aggregate = length)
colnames(data.species) <- sub(". ","_",fixed = TRUE, colnames(data.species))
head(data.species)
data.species <- data.species[,c(-2)]

m.species <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species, chains=4, cores = 1)

plot(m.species)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species)
precis(m.species)
plot(precis(m.species))

# compare the 2 models
compare(m.trt,m.species)
# species model ranked first, weight of 1!
plot(compare(m.trt,m.species))
coeftab(m.trt,m.species)
plot(coeftab(m.trt,m.species))


```

Now model the effect of interaction beween trt and species
```{r}
head(hy.data)
data.species.trt <- hy.data[,c("hyp","species","trt2")]
data.species.trt$id <- 1:nrow(data.species.trt)
data.species.trt <- dcast(data.species.trt, hyp + trt2 + id ~ species,     value.var="species", fun.aggregate = length)
colnames(data.species.trt) <- sub(". ","_",fixed = TRUE, colnames(data.species.trt))
head(data.species.trt)
data.species.trt <- data.species.trt[,c(-3)]

m.species.trt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt2 + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer, bT) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt, chains=4, cores = 1)

# 33 is the mean of the dataset overall

plot(m.species.trt)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt)
precis(m.species.trt)
plot(precis(m.species.trt))

```

Now compare all 3 models
```{r}
compare(m.trt,m.species,m.species.trt) # really likes interaction term
plot(compare(m.trt,m.species,m.species.trt))
coeftab(m.trt,m.species,m.species.trt)
plot(coeftab(m.trt,m.species,m.species.trt))

```

This in and of itself is an indication that trt effects hypocotyl length.
Address this more formally
Looking more closely, the 95% confidence intervals for bT are way above 0:
```{r}
precis(m.species.trt,prob = .95)
# And none of the posterior samples for bT are less than or equal to zero.

bT.post <- extract.samples(m.species.trt)$bT
dens(bT.post, show.zero = TRUE, show.HPDI = 0.95)
sum(bT.post<=0) / length(bT.post) * 100
# none are less than or equal to zero
```

Next:  Repeat with hypocotyl data transformed by square root
```{r}

head(data.species.trt)
data.species.trt.sqrt <- within(data.species.trt, hyp <- sqrt(hyp))

head(data.species.trt.sqrt)

m.species.trt.sqrt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt2 + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=5.7,sd=10),
    c(bChi, bHab, bPen, bPer, bT) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt.sqrt, chains=4, cores = 1)

plot(m.species.trt.sqrt)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt.sqrt)
precis(m.species.trt.sqrt)
round(precis(m.species.trt.sqrt)@output[,1:4]^2,3)
precis(m.species.trt)
plot(precis(m.species.trt.sqrt))

```

Now run model with interaction terms to address treatment by species

```{r}

head(data.species.trt)

m.species.trt.int <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + 
      bT * trt2 + 
      bChi * S_chilense + 
      bHab * S_habrochaites + 
      bPen * S_pennellii + 
      bPer * S_peruvianum +
      bChi_T * S_chilense * trt2 + 
      bHab_T * S_habrochaites * trt2 + 
      bPen_T * S_pennellii * trt2 + 
      bPer_T * S_peruvianum * trt2
      ,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer, bT, bChi_T, bHab_T, bPen_T, bPer_T) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt, chains=4, cores = 1)
# note that chm is missing; so all are being compared to it

# inspect the output
plot(m.species.trt.int)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt.int)
precis(m.species.trt.int)
plot(precis(m.species.trt.int))
# interesting - looks like Hab not different from group effect of trt

```

now compare to the previous models
```{r}
compare(m.trt,m.species,m.species.trt,m.species.trt.int) # nice!
plot(compare(m.trt,m.species,m.species.trt,m.species.trt.int))
coeftab(m.trt,m.species,m.species.trt,m.species.trt.int)
plot(coeftab(m.trt,m.species,m.species.trt,m.species.trt.int))
```
although much more weight given to the interaction model, the dWAIC is small-ish compared to dSE

Below are goals for 8/30/2016

2a) Redo the tomato hyp analysis (or at least the final model with species and treatment) but using square root transformation.  Are model fits improved?  Does it change the parameter estimates?

# first, make new dataframe with sqrt transformed data
```{r}

# modify the dataset; then run below with changed data
m.species.trt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt2 + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer, bT) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt, chains=4, cores = 1)
head(data.species.trt)

```
# how well does model predict data?  
# sample from posterior to get distribution of parameter estimates;
# use the link function!  Page 104
# then find mean of each column to get prediction for each data point (according to model)
# then test correlation wiht actual data; extracting dta from model fit
not a great correlation - was only 0.42 for the original data
for the sqrt transformed data, (after transforming back out into untransformed data!!) and get same correlation

2b) Is WAIC an appropriate way to compare the model fits with untransformed and transformed data?  How can you test this?

2c) How do you get sqrt transformed parameter estimates (especially when they are < 1) back on the original scale (I don't know the answer to this)

3)  Fit a model with a species by treatment interaction.  Which species have a trt response that is different from S. chm?

