---
output:
  html_document:
    keep_md: yes
---
# Statistical Rethinking Chapter 3 problems

__Name:Stacey Harmer__

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 ) 
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
simplehist(likelihood)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
simplehist(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
library(rethinking)
dens(samples)
```

## 3E1
#### How much posterior probability lies below p = 0:2?

```{r}
sum( samples < 0.2 ) / 1e4
```
not much! only 5 e-04

## 3E2
#### How much posterior probability lies above p = 0:8?

```{r}
sum( samples > 0.8) / 1e4
```
0.1117

## 3E3
#### How much posterior probability lies between p = 0:2 and p = 0:8?
```{r}
sum( samples > 0.2 & samples < 0.8 ) / 1e4
```
0.8878

## 3E4
#### 20% of the posterior probability lies below which value of p?
```{r}
quantile(samples, 0.2)
```
0.5195195 

## 3E5
#### 20% of the posterior probability lies above which value of p?
```{r}
quantile(samples, 0.8)
```
0.7567568

## 3E6
#### Which values of p contain the narrowest interval equal to 66% of the posterior probability?
```{r}
HPDI(samples, 0.66)
```
    |0.66     0.66| 
0.5205205 0.7847848 

how similar is this to the PI?
```{r}
PI(samples, 0.66)
```
      17%       83% 
0.5005005 0.7687688 
 
 pretty close.

## 3E7
#### Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?
As above.
```{r}
PI(samples, 0.66)
```
      17%       83% 
0.5005005 0.7687688 

## 3M1
#### Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.

```{r}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )
# define prior
prior <- rep( 1 , 100 )
# compute likelihood at each value in grid
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b")
```


## 3M2
#### Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for p.

```{r}
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
dens(samples)

HPDI(samples, prob = 0.9)
```
    |0.9      0.9| 
0.3535354 0.7272727 

## 3M3
####  Construct a posterior predictive check for this model and data.  Simulate the distribution of samples, averaging over the posterior uncertainty in p.
```{r}
# following example in code 3.11
samples <- sample(p_grid, size = 1e4, replace =T, prob = posterior)
# now generate simulated predictions of observations for 15 tosses
w <- rbinom(1e4, size=15, prob = samples)
simplehist(w)
## 8 is the peak here, but how to determine probability?
sum(w==8)/1e4
# 0.1472
# is that the right way to do it?
```

## 3M4
####  make a new posterior distribution from the 8/15 data.  now calculate the probability of seeing 6 water in 9 tosses
```{r}
# this is the same as above, but using 9 tosses.
x <- rbinom(1e4, size=9, prob = samples)
simplehist(x)
## 5 is the peak here, but how to determine probability?
sum(x==5)/1e4
# 0.2072

```

## 3M5
####  M1, but this time use prior 0 below 0.5 and constant above 0.5.  Then go through the above problems.
```{r}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )
# define prior
prior.2 <- ifelse(p_grid <0.5, 0, 1)
# compute likelihood at each value in grid
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior.2 <- likelihood * prior.2
# standardize the posterior, so it sums to 1
posterior.2 <- unstd.posterior.2 / sum(unstd.posterior.2)

plot(p_grid, posterior.2, type = "b")

# part 2: draw samples from this grid approximation, calculate HPDI
samples.2 <- sample( p_grid , prob=posterior.2 , size=1e4 , replace=TRUE )
dens(samples.2)

HPDI(samples.2, prob = 0.9)
#     |0.9      0.9| 
#  0.5050505 0.7171717 

# part 3 - construct posterior predictive check
w.2 <- rbinom(1e4, size=15, prob = samples.2)
simplehist(w.2)
## 9 is the peak here; determine prob of 9
sum(w==9)/1e4
# 0.1409
# pretty similar to first round

# part 4 - make new posterior distribution; now calculate probability of 6 water in 9 tosses

x.2 <- rbinom(1e4, size=9, prob = samples.2)
simplehist(x.2)
## 6 is the peak here, but how to determine probability?
sum(x.2==6)/1e4
# 0.2387
# now more probable 
```


```{r}
# data for problems
birth1 <- c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
1,0,1,1,1,0,1,1,1,1)
birth2 <- c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
0,0,0,1,1,1,0,0,0,0)
#or, could ahve done this:
library(rethinking)
data(homeworkch3)

```
## 3H1
####  Using grid approximation, compute the posterior distribution for hte probability of a birth being a boy, assuming uniform prior.
```{r}
# I assume this means any birth being a boy; there are 111 boys out of 200 births
p_grid_birth <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep(1, 1000)
likelihood.birth <- dbinom(111, size = 200, prob = p_grid_birth)
simplehist(likelihood.birth)
posterior.birth <- likelihood.birth*prior
posterior.birth <- posterior.birth/sum(posterior.birth)

# what parameter value maxmizes the posterior probability?

p_grid[which.max(posterior.birth)]
#  0.5545546

# does this make sense?  generate some samples and graph out.
samples.birth <- sample( p_grid , prob=posterior.birth , size=1e4 , replace=TRUE )
dens(samples.birth)
max(samples.birth)
# yes, makes sense

```

## 3H2
####  Draw 10,000 random parameter values from the posterior distribution.  then estimate the 50%, 89%, and 97% highest posterior density intervals
```{r}
#  generate some samples from the posterior distribution
samples.birth <- sample( p_grid , prob=posterior.birth , size=1e4 , replace=TRUE )

# HPDI calculations
HPDI(samples.birth, prob = .5)
#      |0.5      0.5| 
# 0.5305305 0.5765766     pretty narrow interval!
HPDI(samples.birth, prob = .89)
#    |0.89     0.89| 
#0.4994995 0.6106106

HPDI(samples.birth, prob = .97)
#     |0.97     0.97| 
# 0.4824825 0.6326326 

## looks good
```

## 3H3
#### Use rbinom to simulate 10,000 replicates of 200 births.  compare this to the real outcome (111 boys per 200 births).  visualize using dens() command.   Does th model fit the data well?
```{r}
#  sample using rbinom and my posterior distribution model (meaning, using my simulated samples generated using my posterior probability)
sim.births <- rbinom( 1e4 , size=200, prob=samples.birth)
dens(sim.births)
# this seems reasonable
## what is HPDI for 90%?

HPDI(sim.births, prob = .90)
# |0.9 0.9| 
#  93  125 
# maybe better:
chainmode(sim.births, adj=0.01)
# 109.0131 
mean(sim.births) # 110.79
median(sim.births) #111
# so yes, this model fits the data very well.  

```

## 3H4
#### now compare 10,000 counts of boys from 100 simulated births only to the number of boys in birth1.
```{r}
# number of first births that are boys
sum(birth1) #51 boys, out of 100 births
# use my model, compare to these data.

# not sure this is correct:
sim.births <- rbinom( 1e4 , size=100, prob=samples.birth)
dens(sim.births)
# 
mean(sim.births) # 55.4
median(sim.births) #55
## this doesn't look like a good fit to these data

```
## 3H5
#### model assumes 1st and 2nd births independent.  is this true?

```{r}
# compare 1e4 simulated counts boys to only those second births that followed girls.  

# count the number of first borns that were girls.  This is 49 
# Simulate 49 births, 10k times.

#  
likelihood.birth.girl.1 <- dbinom(49, size = 100, prob = p_grid_birth)
simplehist(likelihood.birth.girl.1)
posterior.birth.girl.1 <- likelihood.birth.girl.1*prior
posterior.birth.girl.1 <- posterior.birth.girl.1/sum(posterior.birth.girl.1)

samples.birth.girls.1 <- sample( p_grid , prob=posterior.birth.girl.1 , size=1e4 , replace=TRUE )

dens(samples.birth.girls.1)

## no time to finish this, unfortunately.  Not quite right.

```